Modularizing Street Map Generation \\ \vspace{8pt} \large An Adventure in Tensors

### 2013-12-12 paper/paper.tex
added chi^2 formula and more explanation, added more latex processing files to gitignore
added assessment section from google doc

\section{Implementation Details}

\subsection{User-Generated Constraints}

@@

\paragraph{Eigenvectors}
The eigenvalues of a tensor $t$ represented by the matrix $\sqmat{a}{b}{b}{-a}$
are:
\[
    \lambda = \pm\sqrt{a^2+b^2} \\
\]
The major eigenvalue, $\lambda_1$, is the larger value, and the minor
eigenvalue, $\lambda_2$ is the smaller value. The corresponding eigenvectors
are then
\[
    \epsilon_i = \left\{1,\frac{\lambda_i-a}{b}\right\}
\]
The eigenvectors are defined if the tensor field is not degenerate (that is,
$a$ and $b$ are both zero) at that point.

@@

However, even choosing well-spaced seed points will not necessarily result in
an even distribution of streamlines, since two well-spaced seed points can
result in two streamlines which quickly converge. Therefore, we introduce a
third method. To assess how evenly spaced a set of streamlines is, we will
generate a set of discrete points evenly spaced along each streamline, place
these points into $n\times n$ equal-sized buckets covering the bounds of the
space, and perform a chi-squared goodness of fit test on distribution of points
across bucket, where the expected distribution is uniform:

\todo{chi-squared equation}

@@

\section{Assessment of our Work}

@@

\onecolumn
\bibliographystyle{plain}
\bibliography{references}

>>>>

\subsection{User-Generated Constraints}
We begin with a JSON parser, which accepts linear and radial constraints.
Constraints must be input with their type, linear or radial, in addition to
their position in the domain (in the form of x and y coordinates), and finally,
if the constraint is linear, it’s direction in radians and magnitude. In
addition, the executable will accept width and height parameters representing
the size of the domain to build street maps on. Together, the JSON constraints
and width/height parameters form the initial input data that will next be used
to form tensor fields.

@@

\subsection{Eigenvectors}
The eigenvalues of a tensor $t$ represented by the matrix $\sqmat{a}{b}{b}{-a}$
are:
\[
    \lambda = \pm\sqrt{a^2+b^2} \\
\]
The major eigenvalue, $\lambda_1$, is the larger value, and the minor
eigenvalue, $\lambda_2$ is the smaller value. The corresponding eigenvectors
are then
\[
    \epsilon_i = \left\{1,\frac{\lambda_i-a}{b}\right\}
\]
The eigenvectors are defined if the tensor field is not degenerate (that is,
$a$ and $b$ are both zero) at that point.

@@

However, even choosing well-spaced seed points will not necessarily result in
an even distribution of streamlines, since two well-spaced seed points can
result in two streamlines which quickly converge. Therefore, we introduce a
third method. To assess how evenly spaced a set of streamlines is, we will
generate a set of discrete points evenly spaced along each streamline, place
these points into $n\times n$ equal-sized buckets covering the bounds of the
space, and perform a chi-squared goodness of fit test on distribution of points
across bucket, where the expected distribution is uniform:
\[
    \chi^2 = \sum_{i=1}^{n^2} \frac{(O_i-E_i)^2}{E_i}
\]
where $n$ is the number of buckets along each axis, $O_i$ is the observed
number of points in the $i$th bucket, and $E_i=\frac{m}{n^2}$ is the
expected number of points in each bucket, assuming a uniform distribution,
where $m$ is the total number of points in the space.

@@

\section{Assessment}
There were two main goals that we set out to achieve with this project. Here we
present a mostly heuristic, slightly quantitative attempt to qualify our
success, using both internal evaluation and comparisons to existing systems
\cite{chen}.

\subsection{Modularity \& Extensibility}
We believe that we were successful in creating a modular system with reusable
parts that is easy to understand, extend, and modify. One way to illustrate
this is by looking at our pipeline. As a refresher, look at Fig. ***. Each
step of the pipeline has a clear correspondence in the modules that we have
written. We have two modules to deal with constraints, \texttt{Constraint}. To transform
constraints to maps, we have written \texttt{TensorField} and \texttt{Tensor} modules, the latter
of which performs operations on tensors including eigenvector calculations, as
well as \texttt{Streamline}, which has the ability to calculate seed points and trace
lines.

We also created several generally useful modules, including \texttt{Vector2},
which performs basic operations on 2-dimensional vectors and
\texttt{NearestNeighbor}, which defines a type class for a structure that can
store \texttt{Vector2}s and query for the nearest neighbor, as well as a
bucket-based implementation of such a data structure. To deal with input and
output, we created a \texttt{JSONParser} so that our input can come from any
source that is capable of producing JSON, and an \texttt{SVGWriter}, which
deals with the widely-used file format that produces the output seen in many of
the figures in this paper.

In an attempt to compare this with the original work, we consulted the header
files available with Chen’s code. The system has significantly more
functionality than ours, but we had some difficulty discerning which modules
were parallel to ours. The \texttt{tensoranalysis} contains methods for tensor
line tracing, and \texttt{ClosedStreamlineTracing} seems to be related,
although we did not find documentation on the difference between open and
closed tracing. \texttt{SeedPointCtrlDlg} is a dialogue probably related to the
OpenGL framework, but didn’t have the actual calculations of seed points. At
the least, our implementation has a more direct pipeline-to-module correlation,
which will be beneficial for future contributors who wish to build off of our
work.

\subsection{Natural Expression of Underlying Mathematics}
Overall, this seemed to be successful. In general, functional languages are
well suited to express mathematics. Our decision to implement this program in
Haskell meant that that we were able to take advantage of this natural
proclivity towards math in order to make our code easy to understand.

@@

\bibliographystyle{plain}
\bibliography{references}
